\documentclass[11pt]{article}
\title{Problem Set 11}
\input{preamble.tex}

\begin{document} 
  
\section*{Chapter 11}

\begin{enumerate}
  \item Suppose that a monopolist operates in an industry where the demand curve is given by $Q_D = 1000 - 20P$. The firm's marginal cost is constant at $\$ 10$.
  \begin{enumerate}
    \item[(a)] Find the monopolist's marginal revenue curve. (\emph{Note:} it will be easier if you rewrite the market demand curve with $P$ as a function of $Q$).

    \item[(b)] Find the monopolist's profit maximizing level of output.

    \item[(c)] Find the monopolist's profit maximizing price.

    \item[(d)] Find the consumer and producer surplus in this market.
  \end{enumerate}

  \item Assume that a monopolist sells a product with a total cost function $TC = 1200 + 0.5 Q^2$ and a corresponding marginal cost function $MC = Q$. The market demand curve is given by $P = 300 - Q$

  \begin{enumerate}
    \item[(a)] Find the monopolist's marginal revenue curve.

    \item[(b)] Find the monopolist's profit maximizing level of output.

    \item[(c)] Find the monopolist's profit maximizing price.

    \item[(d)] Find the monopolist's marginal cost at the profit-maximizing level of output.

    \item[(e)] Find the price elasticity of demand for this market at the profit-maximizing price and quantity. Does the inverse elasticity pricing rule hold? (Remember that price elasticity of demand is $\frac{\partial Q}{\partial P} \frac{P}{Q}$.)
  \end{enumerate}


  \item Suppose that a single-price monopolist faces the demand curve $P = 100 - Q + \frac{1}{10} I$, where $I$ is average weekly household income, and that the firm's marginal cost function is given by $MC(Q) = 2Q$

  \begin{enumerate}
    \item[(a)] If the average weekly household income is $\$ 600$, then what is the firm's profit-maximizing quantity of output? At what price will the firm sell that output?

    \item[(b)] What will be the consumer and producer surplus in the market?

    \item[(c)] Now suppose that average weekly household income rises by $\$ 400$. Find the new profitmaximizing quantity and price for the monopolist.

    \item[(d)] Find the consumer and producer surplus in the market following the increase in household income.

    \item[(e)] Finally, suppose that, as a result of higher wages in the area, the firm's marginal cost of production has increased to $MC = 3Q$. Taking into account both the increased marginal cost, and the increase in average household income, find the new profit-maximizing price and quantity.

    \item[(f)] Find the new consumer and producer surplus in the market. 
  \end{enumerate}
\end{enumerate}
\end{document}
